---
title: "Untitled"
author: "Andrea Corral Lou"
date: "2023-10-27"
output: html_document
---

(PAQUETES NECESARIOS)

```{r}

library(vegan)
library(permute)
library(lattice)
library(phyloseq)
library(here)
library(tidyr)
library(here)
library(tidyverse)
library(vegan)
library(reshape2)
library(readr)
library(dplyr)
library(Biostrings)
library(phylotools)
library(BiocManager)
library(paletteer)
library(tidyverse)
library(insect)
library(taxonomizr)
library(here)
library(ape)
library(agricolae)
library(Hmisc)
library(remotes)
library(sos)
library(Hmisc)
library(microbiome)  


```

(1) **Voy a crear los tres elementos necesarios para poder usar el formato de phyloseq. Para ello necesito tabla de ASVs, taxa y metadata**

```{r}

# 1. ASVs

ASV.table <- read.csv(here("Outputs/good_output/New_70_90/Supplementary_Table_2_OTU_table.csv"))

#Voy a hacer que cada especies se un OTU, porque sino estoy sobreestimando por la diversidad intraespecifica

OTU <- ASV.table %>% 
  distinct(assign) %>% 
  mutate(numero = row_number()) %>% 
  mutate(OTU = "OTU") %>% 
  mutate(OTU = paste(OTU,numero, sep="")) %>% 
  select(assign, OTU)

ASV.table <- full_join(ASV.table,OTU) %>%
    mutate(species = gsub("\\(species\\)", "", assign),
         species = gsub("\\(genus\\)", "sp.", species),
         species = gsub("\\(family\\)", "sp.", species),
         species = gsub("\\(order\\)", "sp.", species)) 

#ASV <- ASV.table %>% 
#  group_by(Tube, OTU) %>% 
#  dplyr::summarize(nReads=sum(nReads)) %>% 
#  dplyr::rename(Sample = Tube) %>% 
#  pivot_wider(names_from = Sample, values_from = nReads, values_fill = 0) %>% 
#  as.matrix() 
  
#1.1.Vamos a hacer una trasofrmacion de los datos para apalear el sesgo de numero de lecturas y que resulte comparable cuando hacemos beta-diversity

#ASV.table <- read.csv(here("Outputs/good_output/New_70_90/ASV_table_samples_metazoa_without_pos_host_modified.csv"))


```

PAIRWISE

```{r}

a <- ASV.table %>% 
  mutate(adonis = paste(Region, V_o_P, sep="_")) %>% 
  filter(adonis == "Ri_V" | adonis == "Urr_V") %>% 
  select(-adonis)
b <- ASV.table %>% 
  mutate(adonis = paste(Region, V_o_P, sep="_")) %>% 
  filter(adonis == "Ri_V" | adonis == "Ri_P") %>% 
  select(-adonis)
c <- ASV.table %>% 
  mutate(adonis = paste(Region, V_o_P, sep="_")) %>% 
  filter(adonis == "Ri_V" | adonis == "Urr_P") %>% 
  select(-adonis)
d <- ASV.table %>% 
  mutate(adonis = paste(Region, V_o_P, sep="_")) %>% 
  filter(adonis == "Urr_V" | adonis == "Ri_P") %>% 
  select(-adonis)
e <- ASV.table %>% 
  mutate(adonis = paste(Region, V_o_P, sep="_")) %>% 
  filter(adonis == "Urr_V" | adonis == "Urr_P") %>% 
  select(-adonis)
f <- ASV.table %>% 
  mutate(adonis = paste(Region, V_o_P, sep="_")) %>% 
  filter(adonis == "Ri_P" | adonis == "Urr_P") %>% 
  select(-adonis)


ASV <- e %>% #He cambiado Hash por OTU porque para la diversidad son especies no secuencias difernetes
  group_by(Tube, OTU) %>% 
  dplyr::summarize(nReads=sum(nReads)) %>%
  ungroup() %>% 
  select(Tube,OTU,nReads) %>% 
  dplyr::rename(Sample = Tube) %>%
  group_by(Sample) %>% 
  mutate(Total_Sample = sum(nReads)) %>%
  ungroup() %>% 
  mutate(Proportion = (nReads*100)/Total_Sample) %>% 
  select(Sample,OTU,Proportion) %>%
  pivot_wider(names_from = Sample, values_from = Proportion, values_fill = 0) %>% 
  as.matrix() 

ASV.table <- e

```

TRANSFORMACIONES

```{r}

ASV <- ASV.table %>% #He cambiado Hash por OTU porque para la diversidad son especies no secuencias difernetes
  group_by(Tube, OTU) %>% 
  dplyr::summarize(nReads=sum(nReads)) %>%
  ungroup() %>% 
  select(Tube,OTU,nReads) %>% 
  dplyr::rename(Sample = Tube) %>%
  pivot_wider(names_from = Sample, values_from = nReads, values_fill = 0) %>% 
  as.matrix() 


##################################################################################################################
# TRANSFORMACION 1. TENIENDO EN CUENTA ABUNDANCIAS RELATIVAS (TODAS SOBRE 100)
##################################################################################################################

ASV <- ASV.table %>% #He cambiado Hash por OTU porque para la diversidad son especies no secuencias difernetes
  group_by(Tube, OTU) %>% 
  dplyr::summarize(nReads=sum(nReads)) %>%
  ungroup() %>% 
  select(Tube,OTU,nReads) %>% 
  dplyr::rename(Sample = Tube) %>%
  group_by(Sample) %>% 
  mutate(Total_Sample = sum(nReads)) %>%
  ungroup() %>% 
  mutate(Proportion = (nReads*100)/Total_Sample) %>% 
  select(Sample,OTU,Proportion) %>%
  pivot_wider(names_from = Sample, values_from = Proportion, values_fill = 0) %>% 
  as.matrix() 

##################################################################################################################
# TRANSFORMACION 2. TENIENDO EN CUENTA SOLO PRESENCIAS Y AUSENCIAS (0-1) PARA JACCARD
##################################################################################################################

ASV <- ASV.table %>% #He cambiado Hash por OTU porque para la diversidad son especies no secuencias difernetes
  group_by(Tube, OTU) %>% 
  dplyr::summarize(nReads=sum(nReads)) %>%
  mutate(across(nReads, ~if_else(. > 0, 1, 0))) %>% 
  dplyr::rename(Sample = Tube) %>%
  pivot_wider(names_from = Sample, values_from = nReads, values_fill = 0) %>%
  as.matrix() 

##################################################################################################################
# TRANSFORMACION 3. TENIENDO EN CUENTA LA PROP SEGUN EL NUMERO MAXIMO DE LECTURAS OBTENIDAS PARA CADA OTU
##################################################################################################################

ASV <- ASV.table %>% #He cambiado Hash por OTU porque para la diversidad son especies no secuencias difernetes
  group_by(Tube, OTU) %>% 
  dplyr::summarize(nReads=sum(nReads)) %>% 
  dplyr::rename(Sample = Tube) %>%
  group_by(Sample) %>% 
  mutate(Total_Sample = sum(nReads)) %>% 
  mutate(Percentage = (nReads*100)/Total_Sample) %>% 
  ungroup() %>% 
  group_by(OTU) %>% 
  mutate(max_number= max(Percentage)) %>% 
  ungroup() %>% 
  mutate(nReads_transf=(Percentage/max_number)) %>% 
  select(Sample,OTU,nReads_transf) %>%
  mutate(nReads_transf=(nReads_transf*100)) %>% 
  pivot_wider(names_from = Sample, values_from = nReads_transf, values_fill = 0) %>% 
  as.matrix() 

##################################################################################################################
# TRANSFORMACION 4. RAIZ CUADRADA DE LA PROPORCION
##################################################################################################################

ASV <- ASV.table %>% #He cambiado Hash por OTU porque para la diversidad son especies no secuencias difernetes
  group_by(Tube, OTU) %>% 
  dplyr::summarize(nReads=sum(nReads)) %>% 
  dplyr::rename(Sample = Tube) %>%
  group_by(Sample) %>% 
  mutate(Total_Sample = sum(nReads)) %>% 
  mutate(Proportion = (nReads*100)/Total_Sample) %>% 
  ungroup() %>% 
  mutate(Raiz = sqrt(Proportion)) %>% 
  select(Sample,OTU,Raiz) %>%
  pivot_wider(names_from = Sample, values_from = Raiz, values_fill = 0) %>% 
  as.matrix() 

##################################################################################################################
# TRANSFORMACION 5. log-2
##################################################################################################################

ASV <- ASV.table %>% #He cambiado Hash por OTU porque para la diversidad son especies no secuencias difernetes
  group_by(Tube, OTU) %>% 
  dplyr::summarize(nReads=sum(nReads)) %>% 
  dplyr::rename(Sample = Tube) %>%
  group_by(Sample) %>% 
  mutate(Total_Sample = sum(nReads)) %>% 
  mutate(Proportion = (nReads*100)/Total_Sample) %>% 
  ungroup() %>% 
  mutate(Log2 = log(Proportion, base = 2)) %>% 
  select(Sample,OTU,Log2) %>%
  pivot_wider(names_from = Sample, values_from = Log2, values_fill = 0) %>% 
  as.matrix() 


```

CONTINUAMOS

```{r}
#ASV <- ASV.table %>%
#  select(Sample, Hash, nReads) %>%
#  pivot_wider(names_from = Sample, values_from = nReads, values_fill = 0) %>% 
#  as.matrix() 

rownames(ASV) <- ASV[, "OTU"]
ASV = ASV[, -1]
class(ASV) <- "numeric"#"integer"
                              
ASV <- otu_table(ASV, taxa_are_rows = TRUE)

# 2. TAXA

TAXA <- ASV.table %>%
  select(OTU,kingdom,phylum,class,order,family,genus,species) %>% 
  unique() %>% 
  as.matrix() 

rownames(TAXA) <- TAXA[, "OTU"]
TAXA = TAXA[, -1]

#COMBINAMOS ASVs Y TAXA

ASV <- otu_table(ASV, taxa_are_rows = TRUE)
TAXA <- tax_table(TAXA)
physeq = phyloseq(ASV, TAXA)


# 3. METADATA

METADATA <- ASV.table %>% 
  select(Tube,Station,Biol.rep,V_o_P,Region) %>%
  #mutate(Zone = paste (Region,V_o_P, sep="_")) %>% 
  dplyr::rename(Sample = Tube) %>% 
  unique
rownames(METADATA) <- METADATA[, "Sample"]
METADATA = METADATA[, -1] 

#METADATA <- ASV.table %>% 
#  select(Sample,Tube,Station,Biol.rep,rep,V_o_P,Region) %>% 
#  unique
#rownames(METADATA) <- METADATA[, "Sample"]
#METADATA = METADATA[, -1] 

sampledata = sample_data(data.frame(
  Station = METADATA$Station,
  Biol.rep = METADATA$Biol.rep,
  V_o_P = METADATA$V_o_P,
  Region = METADATA$Region,
  row.names=sample_names(physeq),
  stringsAsFactors=FALSE
))


#ADONIS

#METADATA <- ASV.table %>% 
#  semi_join(a, by="Sample") %>% 
#  select(Tube,Station,Biol.rep,V_o_P,Region) %>% 
#  dplyr::rename(Sample = Tube) %>% 
#  unique
#rownames(METADATA) <- METADATA[, "Sample"]
#METADATA = METADATA[, -1] 

#METADATA <- ASV.table %>% 
#  select(Sample,Tube,Station,Biol.rep,rep,V_o_P,Region) %>% 
#  unique
#rownames(METADATA) <- METADATA[, "Sample"]
#METADATA = METADATA[, -1] 

sampledata = sample_data(data.frame(
  Station = METADATA$Station,
  Biol.rep = METADATA$Biol.rep,
  V_o_P = METADATA$V_o_P,
  Region = METADATA$Region,
  row.names=sample_names(physeq),
  stringsAsFactors=FALSE
))


#sampledata = sample_data(data.frame(
#  Tube = METADATA$Tube,
#  Station = METADATA$Station,
#  Biol.rep = METADATA$Biol.rep,
#  rep = METADATA$rep,
#  V_o_P = METADATA$V_o_P,
#  Region = METADATA$Region,
#  row.names=sample_names(physeq),
#  stringsAsFactors=FALSE
#))

#COMBINACION DE LOS TRES ELEMENTOS EN UN OBJETO LLAMADO PHYLOSEQ, TAMBIEN HEMOS AÑADIDO UN ARBOL (NO SE PARA QUE SIRVE)


physeq = merge_phyloseq(physeq, sampledata)

random_tree = rtree(ntaxa(physeq), rooted=TRUE, tip.label=taxa_names(physeq))

physeq = merge_phyloseq(physeq, random_tree)

plot_bar(physeq, fill = "phylum")+
geom_bar(aes(color=phylum, fill=phylum), stat="identity", position="stack") #con esto quitamos el contorno de las barras

plot_bar(physeq, x = "Station", y = "Abundance", fill="phylum") +
geom_bar(aes(color=phylum, fill=phylum), stat="identity", position="stack") #con esto quitamos el contorno de las barras

```

PERMANOVA Y BETADISPER

```{r}

#PERMANOVA

#Generate distance matrix

clr_dist_matrix <- phyloseq::distance(physeq, method = "jaccard") 
sampledf <- data.frame(sample_data(physeq)) %>% 
  mutate(RS = paste(V_o_P, Region, sep ="_"))
vegan::adonis2(clr_dist_matrix ~ V_o_P, data = sampledf)
beta <- betadisper(clr_dist_matrix, sampledf$V_o_P)
permutest(beta)

clr_dist_matrix <- phyloseq::distance(physeq, method = "jaccard") 
sampledf <- data.frame(sample_data(physeq)) %>% 
  mutate(RS = paste(V_o_P, Region, sep ="_"))
vegan::adonis2(clr_dist_matrix ~ Region*V_o_P, data = sampledf)
beta <- betadisper(clr_dist_matrix, sampledf$RS)
permutest(beta)

```

PLOT

```{r}

beta_diversity <- phyloseq::distance(physeq, method = "jaccard") 
beta_pca <- cmdscale(beta_diversity)
plot(beta_diversity, pch = 20)
beta_pcoa <- ordinate(physeq, method = "PCoA", distance = "beta_diversity")
colors <- c("#7fc97f", "#beaed4")

ordination_data <- as.data.frame(beta_pcoa$vectors)
sample_data_df <- data.frame(sample_data(physeq))
ordination_data <- cbind(ordination_data, sample_data_df) %>% 
  mutate(RS = paste0(Region, V_o_P, sep="_"))
centroids <- ordination_data %>%
  group_by(Region) %>%
  dplyr::summarize(PC1 = mean(Axis.1), PC2 = mean(Axis.2))


#PARA SABER QUE VALOR TIENEN LOS PCA
plot_ordination(physeq, beta_pcoa, color = "V_o_P", shape ="Region" ) +
  geom_text(aes(label = Station), hjust = 1, vjust = 1) +
  geom_point(size = 4) +
  scale_color_manual(values = colors)+
  stat_ellipse()


plot_ordination <- ggplot(ordination_data, aes(x = Axis.1, y = Axis.2, color = Region)) +
  geom_point(size = 4) +
  geom_text(aes(label = Station), hjust = 1, vjust = 1) +
  geom_point(data = centroids, aes(x = PC1, y = PC2, color = Region), size = 5, stroke = 2, alpha = 0.7, shape = 18, color = "black") +
  geom_text(data = centroids, aes(x = PC1, y = PC2, label = paste(Region)), vjust = -1, size = 4) +
  scale_color_manual(values = colors) +
  stat_ellipse() +
  labs(title = "Análisis de PCoA con Centroides", x = "PCoA 1", y = "PCoA 2")



ggsave(filename = here("Figures/f_jaccard.pdf"), plot = plot_ordination, width = 7, height = 5, dpi = 300)

```



(2) **Vamos a calcular la diversidad ALFA y vamos a ver si existen diferencias significativas entre diferentes categorias**

```{r}


rich <- estimate_richness(physeq)


#Hay varios indices difernetes con los que puedes calcular la alfa diversity ("Observed", "Chao1", "ACE", "Shannon", "Simpson", "InvSimpson", "Fisher")

alpha_diversity <- estimate_richness(physeq, measures = "Shannon")
hist(alpha_diversity$Shannon)


#Voy a unir esta tabla con la de metadata para poder ver diferentes estadisticos


result <- cbind(sample_data(physeq), alpha_diversity) %>% 
    mutate(loc = case_when(
    str_detect(Station, "Ri1") ~ "Ri_1",
    str_detect(Station, "Ri2") ~ "Ri_2",
    str_detect(Station, "Urr1") ~ "Urr_1",
    str_detect(Station, "Urr2") ~ "Urr_2",
    TRUE ~ Station
  )) %>% 
  mutate(Station=factor(Station,levels = c("Ri1_P","Ri2_P","Ri1_V","Ri2_V","Urr1_P","Urr2_P","Urr1_V","Urr2_V"))) 

alpha.diversity <- ggplot(result, aes(x = Station, y = Shannon)) + 
  geom_boxplot()

ggsave(here("Figures/alpha_diversity_2.pdf"), plot = alpha.diversity, width = 7, height = 5)

#Voy a ver si hay diferencias significativas entre varias cosas:


result <- result %>% 
  mutate(loc_season = paste(loc,V_o_P, sep="_"))

shapiro.test(result$Shannon) #normalidad de los datos. Si el valor p<0.05, los datos no siguen una distribución normal
bartlett.test(result$Shannon,result$loc_season) #homogeniedad de la varianza. Si el valor p<0.05, las varianzas entre los grupos son significativamente diferentes. 

#Una unica variable con un factor de mas de de dos categorias
#anova_result_loc.season <- aov(Shannon ~ loc_season, result)
#summary(anova_result_loc.season)
#TukeyHSD(anova_result_loc.season)
#pairwise.t.test(result$Shannon, result$loc_season, p.adjust.method = "holm")
#
#Una unica variable con dos factores
anova_result_multiple <- aov(Shannon ~ Region*V_o_P, result)
summary(anova_result_multiple)
stats::TukeyHSD(anova_result_multiple)
pairwise.t.test(result$Shannon, result$loc_season, p.adjust.method = "holm")


result_summary <- result %>% 
  group_by(Region,V_o_P) %>% 
  summarise(media=mean(Shannon),
            desviacion=sd(Shannon)
            )

write.csv(result_summary, here("Tables/shannon_2.csv"), row.names = F)


#group_data <- tukey_result$groups[order(rownames(tukey_result$groups)),]
#ggplot(result, aes(x = Station, y = Shannon)) +
#  geom_text(data = data.frame(),
#            aes(x = rownames(group_data), y = max(result$Shannon) + 1, label = group_data$groups),
#            col = 'black',
#            size = 10) +
#  geom_boxplot() +
#  ggtitle("Alpha diversity") +
#  xlab("Site") +
#  ylab("Alpha diversity index")

#plot_richness(physeq, color = "Region", x = "Station", measures = c("Observed", "Chao1", "Shannon")) + geom_boxplot(aes(fill = Station), alpha=.7) + scale_color_manual(values = c('#8dd3c7','#ffffb3','#bebada','#fb8072','#80b1d3','#fdb462','#b3de69','#fccde5')) + scale_fill_manual(values = c('#8dd3c7','#ffffb3','#bebada','#fb8072','#80b1d3','#fdb462','#b3de69','#fccde5'))


#alpha_pd <- estimate_richness(physeq, measures = "Shannon")
#data <- cbind(sample_data(physeq), alpha_pd) 
#psd5.anova <- aov(Shannon ~ Region, data)

```

(3) **Vamos a calcular la diversidad BETA y vamos a hacer una PCA**

```{r}

beta_diversity <- phyloseq::distance(physeq, method = "jaccard") 
beta_pca <- cmdscale(beta_diversity)
plot(beta_diversity, pch = 20)
beta_pcoa <- ordinate(physeq, method = "PCoA", distance = "beta_diversity")
colors <- c("#7fc97f", "#beaed4")

ordination_data <- as.data.frame(beta_pcoa$vectors)
sample_data_df <- data.frame(sample_data(physeq))
ordination_data <- cbind(ordination_data, sample_data_df) 
ordination_data <-ordination_data %>% 
  mutate(RS = paste(Region, V_o_P, sep="_"))
centroids <- ordination_data %>%
  group_by(RS) %>%
  dplyr::summarize(PC1 = mean(Axis.1), PC2 = mean(Axis.2))


#PARA SABER QUE VALOR TIENEN LOS PCA
plot_ordination(physeq, beta_pcoa, color = "V_o_P", shape ="Region" ) +
  geom_text(aes(label = Station), hjust = 1, vjust = 1) +
  geom_point(size = 4) +
  scale_color_manual(values = colors)+
  stat_ellipse()


plot_ordination <- ggplot(ordination_data, aes(x = Axis.1, y = Axis.2, color = RS)) +
  geom_point(size = 4) +
  geom_text(aes(label = Station), hjust = 1, vjust = 1) +
  geom_point(data = centroids, aes(x = PC1, y = PC2, color = RS), size = 3, stroke = 2, alpha = 0.7, shape = 18, color = "black") +
  geom_text(data = centroids, aes(x = PC1, y = PC2, label = paste(RS)), vjust = -1, size = 4) +
  #scale_color_manual(values = colors) +
  stat_ellipse() +
  labs(title = "Análisis de PCoA con Centroides", x = "PCoA 1", y = "PCoA 2") 


#plot_scree(beta_pcoa, "Scree plot for Global Patterns, UniFrac/PCoA")

ggsave(filename = here("Figures/all_jaccard.pdf"), plot = plot_ordination, width = 7, height = 5, dpi = 300)

```




```{r}


beta_diversity <- phyloseq::distance(physeq, method = "bray") 
beta_pca <- cmdscale(beta_diversity)
plot(beta_diversity, pch = 20)
beta_pcoa <- ordinate(physeq, method = "PCoA", distance = "beta_diversity")
colors <- c("#7fc97f", "#beaed4")


plot.ordination <- plot_ordination(physeq, beta_pcoa, color = "V_o_P", shape ="Region" ) +
  geom_text(aes(label = Station), hjust = 1, vjust = 1) +
  geom_point(size = 4) +
  scale_color_manual(values = colors)+
  stat_ellipse()

ordination_data <- as.data.frame(beta_pcoa$vectors)
sample_data_df <- data.frame(sample_data(physeq))
ordination_data <- cbind(ordination_data, sample_data_df)
centroids <- ordination_data %>%
  group_by(V_o_P) %>%
  dplyr::summarize(PC1 = mean(Axis.1), PC2 = mean(Axis.2))

plot_ordination + 
  geom_point(data = centroids, aes(x = PC1, y = PC2, color = V_o_P, shape = Region), size = 5, stroke = 2) +
  geom_text(data = centroids, aes(x = PC1, y = PC2, label = paste(V_o_P, Region)), vjust = -1)


plot_ordination <- ggplot(ordination_data, aes(x = Axis.1, y = Axis.2, color = V_o_P)) +
  geom_point(size = 4) +
  geom_text(aes(label = Station), hjust = 1, vjust = 1) +
  geom_point(data = centroids, aes(x = PC1, y = PC2, color = V_o_P), size = 5, stroke = 2, alpha = 0.7, shape = 2, color = "black") +
  geom_text(data = centroids, aes(x = PC1, y = PC2, label = paste(V_o_P)), vjust = -1, size = 4) +
  scale_color_manual(values = colors) +
  stat_ellipse() +
  labs(title = "Análisis de PCoA con Centroides", x = "PCoA 1", y = "PCoA 2") +
  theme_minimal()


plot_scree(beta_pcoa, "Scree plot for Global Patterns, UniFrac/PCoA")

ggsave(filename = here("Figures/JACCARD_PCOA.pdf"), plot = plot.ordination, width = 7, height = 5, dpi = 300)






plot_heatmap(physeq, method = "NMDS", distance = "bray") #Un heatmap simple



beta_diversity <- phyloseq::distance(physeq, method = "bray") 

#print(beta_diversity)
beta_pca <- cmdscale(beta_diversity)
plot(beta_pca, pch = 20)
beta_pcoa <- ordinate(physeq, method = "PCoA", distance = "beta_diversity")
colors <- c("#7fc97f", "#beaed4")
plot.ordination <- plot_ordination(physeq, beta_pcoa, color = "V_o_P", shape ="Region" ) +
  geom_text(aes(label = Station), hjust = 1, vjust = 1) +
  geom_point(size = 4) +
  scale_color_manual(values = colors)

ggsave(filename = here("Figures/a_bray_PCOA.pdf"), plot = plot.ordination, width = 7, height = 5, dpi = 300)





#Contribucion a la varianza de cada eje

ord_clr <- phyloseq::ordinate(physeq, "NMDS")
phyloseq::plot_scree(ord_clr) + 
  geom_bar(stat="identity", fill = "blue") +
  labs(x = "\nAxis", y = "Proportion of Variance\n")
head(ord_clr$CA$eig)
sapply(ord_clr$CA$eig[1:5], function(x) x / sum(ord_clr$CA$eig))





cap_result <- phyloseq::capscale(beta_pcoa ~ Region + V_o_P, physeq)


ggsave(filename = here("Figures/jaccard_PCoA_pre_aus.pdf"), plot = plot.ordination, width = 7, height = 5, dpi = 300)


#PRUEBAAA

beta_diversity <- phyloseq::distance(physeq, method = "bray") 
vegan::adonis(beta_diversity ~ phyloseq::sample_data(physeq)$Zone)
dispr <- vegan::betadisper(beta_diversity, phyloseq::sample_data(physeq)$Zone)
dispr
plot(dispr, main = "Ordination Centroids and Dispersion Labeled: Aitchison Distance", sub = "")

  
```

**SIMPER: Queremos ver que especies contribuyen mas a que las muestras se separen**

```{r}



#######################################################################################
#
# NMDS PLOT (bray-curtis)
#
#######################################################################################


MM.mds <- metaMDS(MM_spp, distance = "bray", autotransform = FALSE) # Perform nmds
MM.envfit <- envfit(MM.mds, MM_env, permutations = 999) # Fit environmental vectors
MM.sppfit <- envfit(MM.mds, MM_spp, permutations = 999) # Fit species vectors
MM.mds  # Stress value is less than 0.2, which is good.


# Save the results from the nmds and group the data by environmental variables
site.scrs <- as.data.frame(scores(MM.mds, display = "sites"))  

# make depth a grouping variable and save to dataframe
site.scrs <- cbind(site.scrs, Season = MM_env$Season) 

# save NMDS results into dataframe
site.scrs <- cbind(site.scrs, Locality = MM_env$Locality)  

# make temperature a grouping variable and save to dataframe
head(site.scrs)  # View the dataframe



spp.scrs <- as.data.frame(scores(MM.sppfit, display = "vectors"))  # Save species data from nmds to dataframe
spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs))  # Save species values into dataframe
spp.scrs <- cbind(spp.scrs, pval = MM.sppfit$vectors$pvals) # Add species names to dataframe
sig.spp.scrs <- subset(spp.scrs, pval<=0.05) # Add p values to dataframe so you can select species which are significant
head(spp.scrs)  # Show only significant species (using 0.05 as cutoff)



env.scores.MM <- as.data.frame(scores(MM.envfit, display = "vectors"))  # Save environmental variables

# Extract nmds scores of all environmental variables from envifit dataframe
env.scores.MM <- cbind(env.scores.MM, env.variables = rownames(env.scores.MM))  

# Name them 
env.scores.MM <- cbind(env.scores.MM, pval = MM.envfit$vectors$pvals) 

# Add p values to dataframe
sig.env.scrs <- subset(env.scores.barents, pval<=0.05) 

# Show only significant variables (using 0.05 as cutoff)
head(env.scores.barents)  # View dataframe













simp_season <- simper(mat_MM_spp, MM_env$Season)
simp_season
summary(simp_season, ordered = TRUE, digits = 3)

simp_locality <- simper(mat_MM_spp, MM_env$Locality)
simp_locality
summary(simp_locality, ordered = TRUE, digits = 3)

simp_both <- simper(mat_MM_spp, MM_env$Loc_Sea)
simp_both
summary(simp_both, ordered = TRUE, digits = 3)


#EXAMPLE OF SIMPER



barents <- read.csv(here("barents_data.csv"))
barents_spp <- barents %>% select(Re_hi:Tr_spp)
barents_env_raw <- barents %>% select(ID_No:Temperature)
barents_env <- barents_env_raw %>% 
  mutate(Depth_cat=cut(Depth, breaks=c(-Inf, 250, 350, Inf), labels=c("shallow","mid","deep"))) %>%
  mutate(Temp_cat=cut(Temperature, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("vc","c","m","w")))


mat_bar_spp <- as.matrix(barents_spp)
bar_depth <- anosim(mat_bar_spp, barents_env_raw$Depth, distance = "bray", permutations = 9999)

bar_depth 


bar_temp <- anosim(mat_bar_spp, barents_env_raw$Temperature, distance = "bray", permutations = 9999)

bar_temp

simp_depth <- simper(mat_bar_spp, barents_env$Depth_cat)
simp_depth
summary(simp_depth, ordered = TRUE, digits = 3)

simp_temp <- simper(mat_bar_spp, barents_env$Temp_cat)
simp_temp
summary(simp_temp, ordered = TRUE, digits = 3)




barents <- read.csv("barents_data.csv")
# Separate environmental data and species data
barents_spp <- barents %>% select(Re_hi:Tr_spp)

barents_env_raw <- barents %>% select(ID_No:Temperature)


# Create ordinal groups for depth and temperature variables
barents_env <- barents_env_raw %>% 
  mutate(Depth_cat=cut(Depth, breaks=c(-Inf, 250, 350, Inf), labels=c("shallow","mid","deep"))) %>%
  mutate(Temp_cat=cut(Temperature, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("vc","c","m","w")))

# Separate environmental data and species data
barents_spp <- barents %>% select(Re_hi:Tr_spp)

barents_env_raw <- barents %>% select(ID_No:Temperature)


# Create ordinal groups for depth and temperature variables
barents_env <- barents_env_raw %>% 
  mutate(Depth_cat=cut(Depth, breaks=c(-Inf, 250, 350, Inf), labels=c("shallow","mid","deep"))) %>%
  mutate(Temp_cat=cut(Temperature, breaks=c(-Inf, 1, 2, 3, Inf), labels=c("vc","c","m","w")))

# Perform nmds and fit environmental and species vectors
barents.mds <- metaMDS(barents_spp, distance = "bray", autotransform = FALSE)


# Fit environmental vectors
barents.envfit <- envfit(barents.mds, barents_env, permutations = 999) 

# Fit species vectors
barents.sppfit <- envfit(barents.mds, barents_spp, permutations = 999) 

barents.mds

# Save the results from the nmds and group the data by environmental variables
site.scrs <- as.data.frame(scores(barents.mds, display = "sites"))  

# save NMDS results into dataframe
site.scrs <- cbind(site.scrs, Depth = barents_env$Depth_cat)  

# make depth a grouping variable and save to dataframe
site.scrs <- cbind(site.scrs, Temperature = barents_env$Temp_cat) 

# make temperature a grouping variable and save to dataframe
head(site.scrs)  # View the dataframe


# Rename Environmental Factor levels
site.scrs$Temperature <- recode(site.scrs$Temperature, vc = "Very Cold", c = "Cold", m = "Medium", w = "Warm")

site.scrs$Depth <- recode(site.scrs$Depth, shallow = "Shallow", mid = "Mid", deep = "Deep")

# Save species data from nmds to dataframe
spp.scrs <- as.data.frame(scores(barents.sppfit, display = "vectors"))  

# Save species values into dataframe
spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs))  

# Add species names to dataframe
spp.scrs <- cbind(spp.scrs, pval = barents.sppfit$vectors$pvals) 

# Add p values to dataframe so you can select species which are significant
sig.spp.scrs <- subset(spp.scrs, pval<=0.05) 

# Show only significant species (using 0.05 as cutoff)
head(spp.scrs)  # View dataframe
# Save environmental variables
env.scores.barents <- as.data.frame(scores(barents.envfit, display = "vectors"))  

# Extract nmds scores of all environmental variables from envifit dataframe
env.scores.barents <- cbind(env.scores.barents, env.variables = rownames(env.scores.barents))  

# Name them 
env.scores.barents <- cbind(env.scores.barents, pval = barents.envfit$vectors$pvals) 

# Add p values to dataframe
sig.env.scrs <- subset(env.scores.barents, pval<=0.05) 

# Show only significant variables (using 0.05 as cutoff)
head(env.scores.barents)  # View dataframe


# Basic nmds
(nmds.plot.barents <- ggplot(site.scrs, aes(x=NMDS1, y=NMDS2)) + 
# Create the plot
  geom_point(aes(x = NMDS1, y = NMDS2, colour = factor(site.scrs$Depth), shape = factor(site.scrs$Temperature)), size = 2) + 
  # Add site points to plot with the shape determined by temperature and colour determined by depth
  coord_fixed() +
  theme_classic() + 
  theme(panel.background = element_rect(fill = NA, colour = "black", size = 1, linetype = "solid")) +
  labs(colour = "Depth", shape = "Temperature", title = "Does Fish Species Composition Change at\n Varying Water Depths and Temperatures") + 
  # Add legend labels
  theme(legend.position = "right", legend.text = element_text(size = 12), legend.title = element_text(size = 12), axis.text = element_text(size = 10)) # Add legend
)

# Add species vector arrows
(nmds.plot.barents.2 <- nmds.plot.barents +
  geom_segment(data = sig.spp.scrs, aes(x = 0, xend=NMDS1, y=0, yend=NMDS2), arrow = arrow(length = unit(0.25, "cm")), colour = "grey10", lwd=0.3) + 
  # Add vector arrows for significant species
  ggrepel::geom_text_repel(data = sig.spp.scrs, aes(x=NMDS1, y=NMDS2, label = Species), cex = 3, direction = "both", segment.size = 0.25) 
  # Add labels for species (with ggrepel::geom_text_repel so that labels do not overlap)
)

```


PERMANOVA

```{r}

#PERMANOVA

#Generate distance matrix

clr_dist_matrix <- phyloseq::distance(physeq, method = "bray") 
sampledf <- data.frame(sample_data(physeq)) %>% 
  mutate(RS = paste(V_o_P, Region, sep ="_"))
vegan::adonis2(clr_dist_matrix ~ Region * V_o_P, data = sampledf)
beta <- betadisper(clr_dist_matrix, sampledf$RS)
permutest(beta)




#ADONIS test


dispr <- vegan::betadisper(clr_dist_matrix, phyloseq::sample_data(physeq)$Region)
summary(dispr)

#kk <- adonis2(clr_dist_matrix ~ Region * V_o_P, data = phyloseq::sample_data(physeq), permutations = 999)

plot(beta, main = "Ordination Centroids and Dispersion Labeled: Aitchison Distance", sub = "")
boxplot(dispr, main = "", xlab = "")

permutest(dispr)

#Generate distances
ord_unifrac <- ordinate(physeq, method = "PCoA", distance = "wunifrac") 
ord_unifrac_un <- ordinate(physeq, method = "PCoA", distance = "unifrac")   
#Plot ordinations
a <- plot_ordination(physeq, ord_unifrac, color = "Region") + geom_point(size = 2)
b <- plot_ordination(physeq, ord_unifrac_un, color = "Region") + geom_point(size = 2)
cowplot::plot_grid(a, b, nrow = 1, ncol = 2, scale = .9, labels = c("Weighted", "Unweighted"))

# CAP ordinate using Bray Curtis dissimilarity only looking at V4
set.seed(23234) 
ps1V4.rel_bray <- phyloseq::distance(physeq, method = "bray") # CAP ordinate 
cap_ord <- ordinate(physeq = physeq,  
                    method = "CAP", 
                    distance = ps1V4.rel_bray, 
                    formula = ~ Region)

# chech which asix are explaining how mauch variation

scree.cap <- plot_scree(cap_ord, "Scree Plot for MCs in Constrained Analysis of Principal Coordinates (CAPSCALE)")
print(scree.cap)


cap_plot <- plot_ordination(physeq = physeq, 
                            ordination = cap_ord, 
                            color = "Region", 
                            shape = "V_o_P",
                            axes = c(1,2)) + 
  geom_point(size = 3) + 
  geom_point(size = 3) 

cap_plot + ggtitle("CAP_Plot")  + theme_bw()  


# Now add the environmental variables as arrows 
arrowmat <- vegan::scores(cap_ord, display = "bp")
# Add labels, make a data.frame 
arrowdf <- data.frame(labels = rownames(arrowmat), arrowmat)
# Define the arrow aesthetic mapping 
arrow_map <- aes(xend = CAP1, 
                 yend = MDS1, 
                 x = 0, y = 0, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels) 
label_map <- aes(x = 1.3 * CAP1, y = 1.3 * MDS1, 
                 shape = NULL, 
                 color = NULL,
                 label = labels)  

arrowhead = arrow(length = unit(0.02, "npc")) 

##now plot the arrow
cap_plot <- cap_plot + geom_segment(mapping = arrow_map, size = .7, 
                        data = arrowdf, color = "black", 
                        arrow = arrowhead) + 
  geom_text(mapping = label_map, size = 4,  
            data = arrowdf, 
            show.legend = TRUE) + ggtitle("CAP_Plot")  + theme_bw() 



heat.sample <- plot_taxa_heatmap(physeq, subset.top = 10,
    VariableA = "Region",
    heatcolors = brewer.pal(10, "Blues"),
    transformation = "compositional")

```



```{r}

##########
# ANOSIM #
##########

ASV.table <- read.csv(here("Outputs/good_output/New_70_90/ASV_table_samples_metazoa_without_pos_host_modified.csv"))

OTU <- ASV.table %>% 
  distinct(assign) %>% 
  mutate(numero = row_number()) %>% 
  mutate(OTU = "OTU") %>% 
  mutate(OTU = paste(OTU,numero, sep="")) %>% 
  select(assign, OTU)

ASV.table <- full_join(ASV.table,OTU) %>%
    mutate(species = gsub("\\(species\\)", "", assign),
         species = gsub("\\(genus\\)", "sp.", species),
         species = gsub("\\(family\\)", "sp.", species),
         species = gsub("\\(order\\)", "sp.", species)) 



ASV <- ASV.table %>% #He cambiado Hash por OTU porque para la diversidad son especies no secuencias difernetes
  group_by(Tube, OTU) %>% 
  dplyr::summarize(nReads=sum(nReads)) %>% 
  dplyr::rename(Sample = Tube) %>%
  group_by(Sample) %>% 
  mutate(Total_Sample = sum(nReads)) %>% 
  mutate(Percentage = (nReads*100)/Total_Sample) %>% 
  ungroup() %>%
#  mutate(Log2 = log(Percentage, base = 2)) %>% 
#  group_by(OTU) %>% 
#  mutate(max_number= max(Percentage)) %>% 
#  ungroup() %>% 
#  mutate(nReads_transf=(Percentage/max_number)) %>% 
#  select(Sample,OTU,nReads_transf) %>%
#  mutate(nReads_transf=(nReads_transf*100)) %>% 
  pivot_wider(names_from = OTU, values_from = Percentage, values_fill = 0) %>% 
  mutate(Season = case_when(
    str_detect(Sample, "P") ~ "Spring",
    str_detect(Sample, "V") ~ "Summer")) %>% 
  mutate(Locality = case_when(
    str_detect(Sample, "Ri") ~ "Ribera",
    str_detect(Sample, "Urr") ~ "Urrutias")) %>%
  mutate(Loc_Sea = paste(Season, Locality, sep="_")) %>% 
    select(c(1, 79, 80, 81, 4:78))#select(c(77, 78, 2:76)) #
  
MM_spp <- ASV %>% select(c(5:79)) #MM_spp <- ASV %>% select(c(3:77))
MM_env <- ASV %>% select(c(2:3)) #MM_env <- ASV %>% select(c(1:2))
mat_MM_spp <- as.matrix(MM_spp)


MM.mds <- metaMDS(MM_spp, distance = "bray", autotransform = FALSE)
MM.envfit <- envfit(MM.mds, MM_env, permutations = 999)
MM.sppfit <- envfit(MM.mds, MM_spp, permutations = 999)

site.scrs <- as.data.frame(scores(MM.mds, display = "sites"))
site.scrs <- cbind(site.scrs, Locality2 = MM_env$Locality)
site.scrs <- cbind(site.scrs, Season2 = MM_env$Season)
spp.scrs <- as.data.frame(scores(MM.sppfit, display = "vectors"))  
spp.scrs <- cbind(spp.scrs, Species = rownames(spp.scrs)) 
spp.scrs <- cbind(spp.scrs, pval = MM.sppfit$vectors$pvals) 
sig.spp.scrs <- subset(spp.scrs, pval<=0.05)
#env.scores.MM <- as.data.frame(scores(MM.envfit, display = "vectors"))  
#env.scores.MM <- cbind(env.scores.MM, env.variables = rownames(env.scores.MM)) 


(nmds.plot.barents <- ggplot(site.scrs, aes(x=NMDS1, y=NMDS2)) + 
# Create the plot
  geom_point(aes(x = NMDS1, y = NMDS2, colour = factor(site.scrs$Locality2), shape = factor(site.scrs$Season2)), size = 2) + 
  # Add site points to plot with the shape determined by temperature and colour determined by depth
  coord_fixed() +
  theme_classic() + 
  theme(panel.background = element_rect(fill = NA, colour = "black", size = 1, linetype = "solid")) +
  labs(colour = "Depth", shape = "Temperature", title = "Does Fish Species Composition Change at\n Varying Water Depths and Temperatures") + 
  # Add legend labels
  theme(legend.position = "right", legend.text = element_text(size = 12), legend.title = element_text(size = 12), axis.text = element_text(size = 10)) # Add legend
)

# Add species vector arrows
(nmds.plot.barents.2 <- nmds.plot.barents +
  geom_segment(data = sig.spp.scrs, aes(x = 0, xend=NMDS1, y=0, yend=NMDS2), arrow = arrow(length = unit(0.25, "cm")), colour = "grey10", lwd=0.3) + 
  # Add vector arrows for significant species
  ggrepel::geom_text_repel(data = sig.spp.scrs, aes(x=NMDS1, y=NMDS2, label = Species), cex = 3, direction = "both", segment.size = 0.25) 
  # Add labels for species (with ggrepel::geom_text_repel so that labels do not overlap)
)


MM_locality <- anosim(mat_MM_spp, MM_env$Locality, distance = "jaccard", permutations = 9999)
MM_locality

MM_season <- anosim(mat_MM_spp, MM_env$Season, distance = "bray", permutations = 9999)
MM_season

MM_env <- MM_env %>% 
  mutate(Loc_Sea = paste(Locality,Season, sep="_"))

MM_both <- anosim(mat_MM_spp, MM_env$Loc_Sea, distance = "bray", permutations = 9999)
MM_both

#Simper e spara ver que otus son los que están mas asociados a las diferencias detecteadas a tarves de ANOSIM

simp_Locality <- simper(mat_MM_spp, MM_env$Locality)
simp_Locality
summary(simp_Locality, ordered = TRUE, digits = 3)


simp_Season <- simper(mat_MM_spp, MM_env$Season)
simp_Season
summary(simp_Season, ordered = TRUE, digits = 3)


simp_both <- simper(mat_MM_spp, MM_env$Loc_Sea)
simp_both
summary(simp_both, ordered = TRUE, digits = 3)

```



```{r}

##########
# SIMPER #
##########

#SIMPER (Similarity Percentage) is a simple method for assessing which taxa are primarily responsible for an observed difference between groups of samples (Clarke 1993).

## Simper analysis, extract abundance matrix from a phyloseq object
mouse_OTUs = as(otu_table(physeq), "matrix")

# transpose so we have the OTUs as columns
if(taxa_are_rows(physeq)){mouse_OTUs <- t(mouse_OTUs)}

# Coerce the object to a data.frame
mouse_OTUs_scaled = as.data.frame(mouse_OTUs)

# running the simper analysis on the dataframe and the variable of interest "time"
mouse_simper <- simper(mouse_OTUs_scaled, METADATA$Region, permutations = 100)

# printing the top OTUs
print(mouse_simper)
summary(mouse_simper)

```

